PANCHANGE_analysis

### 2020-09-07 PANCHANGE_analysis.Rmd
Added script to create pie charts (ungrouped) for ethnicity, gender, and education. Added a bar plot of age frequencies of whole sample. These are for Kirstin's MQ presentation 08/09/2020

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("broom")
#install.packages("skimr")
#install.packages("polycor")
#install.packages("corrplot")
#install.packages("patchwork")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(polycor)
library(corrplot)
library(patchwork)
library(broom)
library(tidyverse)
```

@@

dat <- dat %>%
  mutate(Ethnicity_collapsed =
           recode_factor(
             Ethnicity_collapsed_numeric,
             `0` = "European",
             `1` = "Mixed or multiple ethnic origins",
             `2` = "Asian or Asian British",
             `3` = "African or African British",
             `4` = "Asian or Asian British"
           )
  )

@@

@@

@@

@@

@@

Create the correlation matrix plot using the corrplot package
```{r Correlation matrix of psychiatric disorders plot, fig.height=7}
corrplot(mhd.corr.matrix, 
         method = "color", # objects to represent the correlations on plot
         type = "lower", # only use the lower triangle of the matrix
         diag = FALSE, # do not show the correlations on the diagonal
         addgrid.col = NA,
         addCoef.col = "black", # colour for the correlation coefficients in the plot
         tl.cex = 0.8,
         tl.col = "black",
         col=colorRampPalette(c("dodgerblue4","white","firebrick4"))(200), # colours for correlations
         # Combine with significance
         p.mat = mhd.p.matrix, # Matrix with p values
         sig.level = 0.01, # Choose significant level
         insig = "blank" # Nonsignificant correlations have no colour
         )
```

>>>>

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("broom")
#install.packages("skimr")
#install.packages("polycor")
#install.packages("corrplot")
#install.packages("patchwork")
#instlal.packages("ggpubr")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(polycor)
library(corrplot)
library(patchwork)
library(broom)
library(tidyverse)
library(ggpubr)
```

@@

dat <- dat %>%
  mutate(Ethnicity_collapsed =
           recode_factor(
             Ethnicity_collapsed_numeric,
             `0` = "European",
             `1` = "Mixed or multiple ethnic origins",
             `2` = "Asian or Asian British",
             `3` = "African or African British",
             `4` = "Other"
           )
  )

@@

age bar plot for whole sample (not grouped)

Create a nice summary dataset for age and display it as a bar plot

```{r age_category prop total (ungrouped)}

age_category_prop_total <- dat %>%
  count(age_category) %>%
  mutate(Prop = round(n/sum(n), 2))

# add a column to help order the factor
age_category_prop_total$order <- c(1,2,3,4,5,6,7,8,9,10,11,12,13)

# total responders
ageN <- sum(age_category_prop_total$n)

kable(age_category_prop_total)
age_category_prop_total

```{r age_category prop total (ungrouped) plot}
#Colour palette
RAMPagepalette <- c(
  "#00D9AC","#21D9B3","#4DD9BC","#78D9C5","#8ED9C9","#A3D9CE",
  "#CFD9D7",
  "#DAD0BA","#DAC38F","#DAC38F","#DAB663" ,"#DAA937","#DA9C0C")

#Bar plot
plot.age <-

  age_category_prop_total %>%
    mutate(age_category = fct_reorder(age_category,order)) %>%

  ggplot(aes(x = age_category)) +
  geom_bar(aes(y = n, color=age_category, fill=age_category, alpha = 1),
           stat ="summary") +
  scale_fill_manual(values =  RAMPagepalette) +
  scale_color_manual(values =  RAMPagepalette) +
  labs(x = "",
       y = "Total count",
       title = paste("Number of participants in each age category",sep=" "),
       subtitle = paste("n = ", ageN,
                        sep = ""),
       fill = "Age category") +
   guides(alpha = FALSE, color = FALSE, fill = FALSE) +
  scale_x_discrete(limits = (levels(age_category_prop_total$age_category))) +
  theme_classic2() +
   theme(axis.text.x = element_text(colour="black", size = 12,angle=45,
                                    vjust=0.5),
        axis.text.y = element_text(colour="black", size = 12))

plot.age

@@

### pie charts

```{r Ethnicity_collapsed prop per total summary}

Ethnicity_collapsed_prop_Sample <- dat %>%
  count(Ethnicity_collapsed) %>%
  drop_na() %>%
  mutate(freq = round(n/sum(n), 2))

# total responders
ethN <- sum(Ethnicity_collapsed_prop_Sample$n)

kable(Ethnicity_collapsed_prop_Sample)
Ethnicity_collapsed_prop_Sample

pie chart with only European label

```{r ethnicity pie chart biggest group label}

eth_pie_alllabel <-

  Ethnicity_collapsed_prop_Sample %>%
   mutate(Ethnicity_collapsed = fct_reorder(Ethnicity_collapsed,-n)) %>%

  ggplot(aes(x="", y=freq, fill=Ethnicity_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette6) +
  scale_color_manual(values =  COPINGpalette6) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported ethnicity",
                      sep=" "),
        subtitle = paste("n =", sum(Ethnicity_collapsed_prop_Sample$n),
                         sep = " "),
        fill = "Self reported ethnicity")  +
    geom_text(aes(y = c(0.5),
                  label= c(paste0(round((freq[1]*100),0),"%"))), #Frequency of European only
             size=5,
             face="bold"
             )

eth_pie_alllabel

pie chart - no labels

```{r ethnicity pie chart no label}

eth_pie <-

    Ethnicity_collapsed_prop_Sample %>%
   mutate(Ethnicity_collapsed = fct_reorder(Ethnicity_collapsed,-n)) %>%

  ggplot(aes(x="", y=freq, fill=Ethnicity_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette6) +
  scale_color_manual(values =  COPINGpalette6) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported ethnicity",
                      sep=" "),
        subtitle = paste("n =", sum(Ethnicity_collapsed_prop_Sample$n),
                         sep = " "),
        fill = "Self reported ethnicity")

eth_pie

@@

```{r summary gender}

summarytools::freq(data_dem[(grep("gender",names(data_dem)))],
                   style = "rmarkdown")

### pie charts

```{r Gender_collapsed prop per total}

Gender_collapsed_prop_total <- dat %>%
  count(Gender_collapsed) %>%
  mutate(freq = round(n/sum(n), 2))

kable(Gender_collapsed_prop_total)
Gender_collapsed_prop_total

```{r gender pie chart - all labels}

gen_pie_alllabel <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")  +
    geom_text(aes(y = c(0.85,0.35,0),
                  x = c(1.2,1.2,1.8),
                  label= c(paste0(round((freq[1:2]*100),0),"%"),paste0(round((freq[3]*100),1),"%"))),
             size=5,
             face="bold"
             )

gen_pie_alllabel

pie chart - only labelling two largest groups (male and female)

```{r gender pie chart - male/female labels only}

gen_pie <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")  +
    geom_text(aes(y = c(0.85,0.35,0),
                  x = c(1.2,1.2,1.8),
                  label= c(paste0(round((freq[1:2]*100),0),"%"),"")),
             size=5,
             face="bold"
             )

gen_pie

pie chart -  no labels

```{r gender pie chart - no labels }

gen_pie_nolabel <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")

gen_pie_nolabel

@@

### pie chart

summary dataset with frequencies

```{r highest education prop per total}

highest_education_prop_total <- dat %>%
  count(highest_education) %>%
  mutate(freq = round(n/sum(n), 2)) %>% 
  drop_na()

kable(highest_education_prop_total)
highest_education_prop_total

# total responders
eduN <- sum(highest_education_prop_total$n)

edubindat <- dat %>%
  mutate(demographics.education_binary =
           case_when(
             highest_education_numeric == "4" ~ "Bachelors or higher",
             highest_education_numeric == "3" ~ "A-levels or lower"
           )) %>%
  summarise(n = n()) %>%
  drop_na() %>%
  mutate(freq = n / sum(n))

educatdat <- dat %>%
  mutate(demographics.education_three = 
           case_when(
             highest_education_numeric == "4" ~ "Bachelors or higher",
             highest_education_numeric == "3" ~ "A-levels",
             highest_education_numeric == "1" ~ "GCSEs or lower"
           )) %>%
  summarise(n = n()) %>%
  drop_na() %>%
  mutate(freq = n / sum(n))

pie chart labels

```{r education pie chart all labels}

edu_pie_alllabel <-

    highest_education_prop_total %>%
    mutate(demographics.highest_level_of_education = fct_reorder(highest_education,-n)) %>%

  ggplot(aes(x="", y=freq, fill=demographics.highest_level_of_education)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette7) +
  scale_color_manual(values =  COPINGpalette7) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported level of education",
                      sep=" "),
        subtitle = paste("n", sum(highest_education_prop_total$n),
                         sep = " "),
        fill = "Self reported level of education")  +
    geom_text(aes(y = c(0.13,0.04,0.3,0.65),
                  x = c(1.2,1.2,1.2,1.2), #GCSE, NVQ, A-levels, university
                  label= paste0(round((freq*100),0),"%")),
             size=5,
             face="bold"
             )

edu_pie_alllabel

#y = c(0.57,0.85,0.25,0.38,0.015,0.05,0.14),
#x = c(1.2,1.2,1.2,1.2,1.35,1.2,1.2)

```{r binary education pie chart all labels}

edu_pie_binlabel <-

    edubindat %>%
    mutate(demographics.education_binary = fct_reorder(demographics.education_binary,-n)) %>%

  ggplot(aes(x="", y=freq, fill=demographics.education_binary)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette7) +
  scale_color_manual(values =  COPINGpalette7) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP sample self reported highest level of education, above or below bachelor degree or equivalent",
                      sep=" "),
        subtitle = paste("n", sum(edubindat$n),
                         sep = " "),
        fill = "Self reported level of education")  +
    geom_text(aes(y = c(0.25,0.75),
                  label= paste0(round((freq*100),0),"%")),
             size=5,
             face="bold"
             )

edu_pie_binlabel

```{r 3 category education pie chart all labels}

edu_pie_catlabel <-

    educatdat %>%
    mutate(demographics.education_three = fct_reorder(demographics.education_three,-n)) %>%

  ggplot(aes(x="", y=freq, fill=demographics.education_three)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette7) +
  scale_color_manual(values =  COPINGpalette7) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP sample self reported highest level of education, above or below bachelor degree or equivalent",
                      sep=" "),
        subtitle = paste("n", sum(educatdat$n),
                         sep = " "),
        fill = "Self reported level of education")  +
    geom_text(aes(y = c(0.25,0.75,0.02),
                  label= c(paste0(round((freq[1:2]*100),0),"%"),"") ),
             size=5,
             face="bold"
             )

edu_pie_catlabel

```{r education pie chart - largest groups labelled }

edu_pie_somelabel <-

    edudat %>%
    mutate(demographics.highest_level_of_education = fct_reorder(demographics.highest_level_of_education,-n)) %>%

  ggplot(aes(x="", y=freq, fill=demographics.highest_level_of_education)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette7) +
  scale_color_manual(values =  COPINGpalette7) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP sample self reported level of education",
                      sep=" "),
        subtitle = paste("n", sum(edudat$n),
                         sep = " "),
        fill = "Self reported level of education")  +
    geom_text(aes(y = c(0.57,0.85,0.25,0.38,0.015,0.05,0.14),
                  x = c(1.2,1.2,1.2,1.2,1.35,1.2,1.2),
                  label=c( paste0(round((freq[1:4]*100),0),"%"),"",
                  paste0(round((freq[6:7]*100),0),"%"))),
             size=5,
             face="bold"
             )
edu_pie_somelabel

pie chart -  no labels

```{r education pie chart no labels }

edu_pie_nolabel <- 
   edudat %>%
    mutate(demographics.highest_level_of_education = fct_reorder(demographics.highest_level_of_education,-n)) %>%

  ggplot(aes(x="", y=freq, fill=demographics.highest_level_of_education)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette7) +
  scale_color_manual(values =  COPINGpalette7) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP sample self reported level of education",
                      sep=" "),
        subtitle = paste("n", sum(edudat$n),
                         sep = " "),
        fill = "Self reported level of education")

edu_pie_nolabel

@@

Create the correlation matrix plot using the corrplot package
```{r Correlation matrix of psychiatric disorders plot, fig.height=7}
corrplot(mhd.corr.matrix, 
         method = "color", # objects to represent the correlations on plot
         type = "lower", # only use the lower triangle of the matrix
         diag = FALSE, # do not show the correlations on the diagonal
         addgrid.col = NA,
         addCoef.col = "black", # colour for the correlation coefficients in the plot
         tl.cex = 0.8,
         tl.col = "black",
         col=colorCOPINGpalette(c("dodgerblue4","white","firebrick4"))(200), # colours for correlations
         # Combine with significance
         p.mat = mhd.p.matrix, # Matrix with p values
         sig.level = 0.01, # Choose significant level
         insig = "blank" # Nonsignificant correlations have no colour
         )
```
